Union and intersection contracts are hard, actually

### 2021-07-26 paper.tex
Hardcode no-editing-marks to make it work with git-latexdiff
Revision: small precisions

Higher-order contracts described in Section~\ref{sub-sec:contracts} already
introduce side-effects: semantically, the \nickel{blame} operator signaling
contract violation is equivalent to non-termination (error reporting aside).
But non-termination is a special kind of effect, if only because it preserves
referential transparency. In a lazy language like Nickel,
(\ref{eq:cbn-expansion}) still holds for all terms in the presence of
non-termination. In a strict language (call-by-value), (\ref{eq:cbn-expansion})
fails in all generality because the transformation may cause a previously
unevaluated term to be suddenly evaluated.  However, (\ref{eq:cbn-expansion})
holds given slight restrictions on the context $M$: for example, it holds as
long as at least one occurrence of $x$ in $M$ is in scope (i.e. not inside a
function). Beside, many languages out there already feature non-termination. For
such langauges, adding contracts does not change what optimizations are valid.

@@

Figure~\ref{fig:optimized-programs} illustrates this point further. It contains
an original program and an optimized version where the common subexpression
\nickel{f 1} has been eliminated. While equivalent in a pure language without
contracts, these two programs behave differently because of unions:

>>>>

Higher-order contracts described in Section~\ref{sub-sec:contracts} already
introduce side-effects: semantically, the \nickel{blame} operator signaling
contract violation is equivalent to non-termination (error reporting aside).
But non-termination is a special kind of effect, if only because it preserves
referential transparency. In a lazy language like Nickel,
(\ref{eq:cbn-expansion}) still holds for all terms in the presence of
non-termination. In a strict language (call-by-value), (\ref{eq:cbn-expansion})
fails in all generality because the transformation may cause a previously
unevaluated term to be suddenly evaluated.  However, (\ref{eq:cbn-expansion})
holds given slight restrictions on the context $M$: for example, it holds as
long as at least one occurrence of $x$ in $M$ is in scope (i.e. not inside a
function). Beside, many languages out there (including Nickel) already feature
non-termination. For such languages, adding contracts does not change what
optimizations are valid.

@@

Figure~\ref{fig:optimized-programs} illustrates this point further. It contains
an original program and an optimized version where the common subexpression
\nickel{f 1} has been eliminated. While equivalent in a pure language with only
non-termination or plain higher-order contracts, these two programs behave
differently because of unions:
